---
title: 'MidTermProject : Pricipal Component Analysis and Multiple Linear Regression'
author: "Smita Sukhadeve"
STA 588
output: word_document
---

#Analsis of original Dataset:
```{r}

ames0 = read.csv("http://people.usm.maine.edu/cpeng/datasets/amescsv.csv")
attach(ames0)
# dimension of the dataset
dim(ames0)
# Summary of the Dataset
summary(ames0)
```

#Observation of NA Values in the dataset
```{R}
# Missing Values in the dataset corresponding to each predictor
na_count <-sapply(ames0, function(x) sum(length(which(is.na(x)))))
na_count
```

#Creation of Work Datasets
```{R}

#Excluded Varaibles
#PID : Nominal
#Following variable removed Due to large amount of missing Values :
#PoolQC, MiscFeature, Alley, Fence, FireplaceQu around 1500 values missing
#LotFrontage : 490 missing
#GarageYrBlt, GarageQual, GarageCond, GarageType, GarageFinish : NAs above 100
#BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2 : around 70
# Working Set With Excluded Variables
#Removed 19 variables: Due to Missing values

workingSet1 = ames0[ , c(-2, -8, -32:-35, -37, -59, -60:-62,  -65, -66, -74:-76) ]

```

# Working Set 2: Excluded DISCRETE Variables from the study
```{R}

# Discrete variables explain additional features such bathroom, YearBuild, yearModeled
# Discrete variable indexes: -49:-54, -56, -58, -61, -63, -78, -79

workingSet2 = ames0[ , c(-2, -8, -32:-35, -37, -59, -60:-62,  -65, -66, -74:-76,
                        -49:-54, -56, -58, -61, -63, -78, -79) ]

```

# Create new subset for the continuous variables from working set2 : Will be used during PCA
```{R}
# Order: Nominal is included for the identification puprpose of a specific record
continuous.var = c (     "Order", "LotFrontage", "LotArea", "MasVnrArea", "BsmtFinSF1", "BsmtFinSF2",
                         "BsmtUnfSF", "TotalBsmtSF", "X1stFlrSF" , "X2ndFlrSF", "LowQualFinSF", 
                         "GrLivArea", "GarageArea", "WoodDeckSF", "OpenPorchSF", "EnclosedPorch", 
                         "X3SsnPorch", "ScreenPorch", "PoolArea", "MiscVal", "SalePrice" )

contVar.dat <- subset(workingSet2, select = continuous.var)
dim(contVar.dat)
```

#Section 3.1 : Full Model With all continuous predictors
```{R}
lm_allCont = lm(SalePrice ~.-Order, data = contVar.dat)
summary(lm_allCont)
par(mfrow =c (1,1))
plot(lm_allCont)
```

#Removed Outliers from the Working Dataset
```{R}
# Full model result after removing Outliers
workingSet2 = workingSet2[which(workingSet2$GrLivArea < 4000), ]
contVar.dat <- subset(workingSet2, select = continuous.var)
dim(contVar.dat)
lm_noOutlier = lm(SalePrice~ .-Order, data = contVar.dat)
summary(lm_noOutlier)
```

#Section 3.2 - Full regression model using continuous predictors with NA's replaced with corresponding Mean values
```{R}
#Replaced NAs by cooresponding mean values
contVar.dat$MasVnrArea[which(is.na(contVar.dat$MasVnrArea))]<- mean(contVar.dat$MasVnrArea, na.rm = TRUE)
contVar.dat$BsmtFinSF1[which(is.na(contVar.dat$BsmtFinSF1))]<- mean(contVar.dat$BsmtFinSF1, na.rm = TRUE)
contVar.dat$BsmtFinSF2[which(is.na(contVar.dat$BsmtFinSF2))]<- mean(contVar.dat$BsmtFinSF2, na.rm = TRUE)
contVar.dat$BsmtUnfSF[which(is.na(contVar.dat$BsmtUnfSF))]<- mean(contVar.dat$BsmtUnfSF, na.rm = TRUE)
contVar.dat$TotalBsmtSF[which(is.na(contVar.dat$TotalBsmtSF))]<- mean(contVar.dat$TotalBsmtSF, na.rm = TRUE)
contVar.dat$GarageArea[which(is.na(contVar.dat$GarageArea))]<- mean(contVar.dat$GarageArea, na.rm = TRUE)
na_count <-sapply(contVar.dat, function(x) sum(length(which(is.na(x)))))
na_count
names(contVar.dat)
#Fitting linear Model
lm_withNoNull = lm(contVar.dat$SalePrice ~.-Order, data = contVar.dat[,-2])
summary(lm_withNoNull )
library(car)
#vif(lm_withNoNull)
```

#Section 2.21 - Exploration of continuous data using pair-wise scatter plots
```{R}
library(car)
# panel.smooth function is built in.
# cor.display puts correlation in upper panels, size proportional to correlation
cor.display <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}

#Scatter plot for few continuos variables
pairs(~LotArea+BsmtUnfSF+X1stFlrSF+X2ndFlrSF+GrLivArea+GarageArea+SalePrice, data=contVar.dat,
      lower.panel=panel.smooth, upper.panel=cor.display, 
      pch=20, main="Ames Scatterplot Matrix")

#Correlation Matrix
corr_Matrix = round(cor(contVar.dat[, c(-1, -20)]), 5)
corr_Matrix[which(corr_Matrix < 0.45)]<- NA
#Density disctribution for the Sale Price
plot(density(ames0$SalePrice), main='Density of SalePrice for Ames',
                         xlab= 'SalePrice', ylab='density')

```


#3.3  Principal Component Analysis 
```{R}

# excluded SalePrice: Response variable and X1stFlrSF, X1stFlrSF
names(contVar.dat)
ames.pc = princomp(contVar.dat[, c(-1, -2, -21, -9, -10)], cor = TRUE)
summary(ames.pc, loadings = TRUE)

require(graphics)

# Two dimensional view of the Data using first two principal components
library(lattice)
screeplot(ames.pc, npcs = 10, type = "lines", main = "Scree Plot of Principal Components" )
biplot (ames.pc , scale =1, col = c(1,3))




#Using first 10 principal componets
amesTranformed.dat = data.frame(ames.pc$scores[, 1:9])


```

# Combining Saleprice and Perform Linear regression modelling
```{R}

PCA_SalePrice_dat = cbind(amesTranformed.dat,  SalePrice= contVar.dat[, 21])
names(contVar.dat)
dim(PCA_SalePrice_dat)
summary(PCA_SalePrice_dat$SalePrice)
```

#Handling Categorical Variables
```{r}

categorical_dat = workingSet2[ , sapply(workingSet2, is.factor)]
str(categorical_dat )
names(categorical_dat)

for (i in c(1:29)) {  
     categorical_dat[,i] <- as.factor(categorical_dat[,i]) 
}
summary(categorical_dat)
dim(categorical_dat)
```

# Combine Continous and categorical variables
```{r}
finalWorkingSet = cbind(PCA_SalePrice_dat, categorical_dat)
dim(finalWorkingSet)
summary(finalWorkingSet)
names(finalWorkingSet)
```

# 3.4 Multiple linear Regression by combining the categorical and Continuous
```{R}

lmFinal.full = lm(SalePrice ~ ., data = finalWorkingSet)
lmFinal.null = lm(SalePrice ~ 1, data = finalWorkingSet)
```

# Section 3.5 Stepwise forward model Selection
```{R}
step(lmFinal.null, scope = list(lower = lmFinal.null, upper = lmFinal.full ), direction = "forward")

forward =  lm(formula = SalePrice ~ Comp.1 + Neighborhood + KitchenQual + 
    ExterQual + HouseStyle + BldgType + MasVnrType + SaleCondition + 
    Exterior1st + Functional + Foundation + Comp.8 + Comp.2 + 
    Condition1 + HeatingQC + LandSlope + RoofMatl + LandContour + 
    Street + Comp.5 + Comp.3 + LotConfig + ExterCond + Condition2 + 
    RoofStyle + Comp.7 + Utilities + Comp.4 + SaleType, data = finalWorkingSet)
summary(forward)
par(mfrow = c (2,2))
plot(forward)
```


#3.6 Stepwise Backward Model Selection
```{R}

step(lmFinal.full, data = finalWorkingSet, direction = "backward")
backward = lm(formula = SalePrice ~ Comp.1 + Comp.2 + Comp.3 + Comp.4 + 
    Comp.5 + Comp.7 + Comp.8 + Street + LandContour + Utilities + 
    LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + 
    BldgType + HouseStyle + RoofStyle + RoofMatl + Exterior1st + 
    MasVnrType + ExterQual + ExterCond + Foundation + HeatingQC + 
    KitchenQual + Functional + SaleType + SaleCondition, data = finalWorkingSet)
summary(backward)
plot(backward)
```


#3.7 Model Improvement By removing insignificant variables
```{R}
forward$coefficients
forward1 = update(forward, ~.-LotConfig)
summary(forward1) 
forward2= update(forward1, ~. -RoofStyle)
summary(forward2)

forward3= update(forward2, ~. -ExterCond)
summary(forward3)

forward4= update(forward3, ~. -Condition2)
summary(forward4)

forward5= update(forward4, ~. -Exterior1st)
summary(forward5)

```

# Combining the Categorical Variable levels 
```{R}

#analysing the frequencies following categorical variable
par(mfrow = c(3, 3))
barplot(prop.table(table(categorical_dat$Functional)), col = 3, xlab = "Functional")
barplot(prop.table(table(categorical_dat$HeatingQC)), col = 3, xlab = "HeatingQC")
barplot(prop.table(table(categorical_dat$Condition1)), col = 3, xlab = "Condition") 
barplot(prop.table(table(categorical_dat$KitchenQual)), col = 3, xlab = "KitchenQual") 

#Comining variable levels
WS_combLevel = finalWorkingSet
levels(WS_combLevel$Functional)<-c("NonTyp", "NonTyp", "NonTyp", "NonTyp", "NonTyp" , "NonTyp" , "NonTyp", "Typ")
levels(WS_combLevel$HeatingQC)<-c("Gd", "Po", "Gd", "Po", "Gd")
levels(WS_combLevel$Condition1)= c("AbNorm", "AbNorm"  , "Norm",    "AbNorm" ,  "AbNorm",   "AbNorm",   "AbNorm",   "AbNorm",   "AbNorm")
levels(WS_combLevel$KitchenQual) = c("Gd", "Po", "Gd", "Po", "Gd")
par(mfrow = c(1,1))
```

```{R}

finalmodel = lm(formula = SalePrice ~ Comp.1 + Neighborhood + KitchenQual + 
    ExterQual + HouseStyle + BldgType + MasVnrType + SaleCondition + 
    Functional + Foundation + Comp.8 + Comp.2 + Condition1 + 
    HeatingQC + LandSlope + RoofMatl + LandContour + Street + 
    Comp.5 + Comp.3 + Comp.7 + Utilities + Comp.4 , 
    data = WS_combLevel)
summary(finalmodel)
plot(finalmodel)
finalmodel$
confint(finalmodel)
vif(finalmodel)
```

